---
title: "Manhattan Analysis"
author: "Nick Grammas"
date: 2020-12-01T21:13:14-05:00
categories: ["R"]
tags: ["R Markdown", "plot", "regression"]
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Introduction:

In this section I look at the borough Manhattan. In this analysis I have created plots that look at square feet vs price, down return vs sale price, sale price vs projected rental income, and projected mortgage payments vs ROI.

```{r echo=FALSE, message=FALSE, results = 'hide'}
library(tidyverse)
library(here)
nyc_properties <- read_csv(here("Data/NYC_property_sales.csv"))
manhattan_properties <- nyc_properties %>% filter(borough == "Manhattan")
```
```{r}
manhattan_analysis <- manhattan_properties %>%
  mutate(avg_price = sum(sale_price) / 6572) %>%
  mutate(projected_rental_income = gross_square_feet * 3.87) %>%
  mutate(down_payment = sale_price * .20) %>%
  mutate(principal = sale_price - down_payment) %>%
  mutate(projected_mortgage_payment = ((principal/330) * 1.042)) %>%
  mutate(projected_mortgage_yr = projected_mortgage_payment * 12) %>%
  mutate(projected_rental_income_yr = (projected_rental_income * 12)) %>%
  mutate(ROI = (projected_rental_income_yr - projected_mortgage_yr) / projected_mortgage_yr) %>%
  mutate(down_return = projected_rental_income_yr / down_payment * 100)
```
```{r echo=FALSE, message=FALSE, results = 'hide'}
library(plotly)
```
```{r}
plot1 <- ggplot(data = manhattan_analysis, aes(x = gross_square_feet, y = sale_price, 
                                               label = address)) +
  geom_point() +
  geom_smooth()

ggplotly(plot1, tooltip = 'label')
```

This plot once again shows that as square feet increases in Manhattan the price tends to increase as well. Unlike most of the other boroughs in this analysis the points in this plot are much more spread out. This may not come as a surprise after looking at the mean property prices for each borough in the NYC General Analysis section. In that section, Manhattan had the highest mean sale price, but also had the largest variability of all the boroughs. This plot seems to support that. 

Down Return Vs Price:

```{r}
plot2 <- ggplot(data = manhattan_analysis, aes(x = down_return, y = sale_price, 
                                               label = address)) +
  geom_point() +
  geom_smooth()

ggplotly(plot2, tooltip = 'label')
```

The down return trend line is very similar to all the boroughs in this analysis, but the returns for Manhattan on lower priced properties is much higher than others. It seems as if Manhattan is certainly the most desirable borough within the city, and therefore the demand to live in this borough could be a major factor in getting these great returns on down payments. 

Price Vs Projected Rental Income:

```{r}
plot3 <- ggplot(data = manhattan_analysis, aes(x = sale_price, y = projected_rental_income_yr, 
                                               label = address)) +
  geom_point() +
  geom_smooth()
ggplotly(plot3, tooltip = 'label')
```

The plot shows that as the price of a property increases the projected rental income tends to increase as well. This is very similar to other boroughs, but there are many properties spread out on this plot, and many that have projected rental incomes much higher than what the trend line would predict. Again, this could be due to the demand from consumers to live in this area of the city. 

Projected Mortgage Payments Vs ROI

```{r}
plot4 <- ggplot(data = manhattan_analysis, aes(x = projected_mortgage_yr, y = ROI, 
                                               label = address)) +
  geom_point() +
  geom_smooth()
ggplotly(plot4, tooltip = 'label') 
```

This plot shows that higher mortgage costs for the year tend to produce very small returns on the investments. The only properties that seem to producing returns are properties that do not have extremely high mortgage costs. That being said, based on some of the other plots and trends in this analysis, the low return may not out weight the costs of owning property in this borough. 

